---
title: "Hw04"
author: "Vetle Birkeland Huglen"
date: "4 10 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(gapminder)
library(tidyverse)
```

##TASK 1 - TIDY DATA

Here follows a table and a graph visualising life expectancy by year in Norway, Canada and Sudan.

```{r}
compareLifeExp <- gapminder %>% 
  select(year, country, lifeExp) %>%
  filter(country == 'Norway' | country == 'Canada' | country == 'Sudan') %>%
  group_by(year) %>%
  spread(key = country, value=lifeExp)
knitr::kable(compareLifeExp)

compareLifeExp %>%
  ggplot() +
  geom_point(aes(year, Norway), colour = 'Orange') +
  geom_point(aes(year, Canada), colour = 'Green') +
  geom_point(aes(year, Sudan), colour = 'Turquoise') +
  labs(y = "Life Expectancy")
```

Now lets look at the smallest life expectancy in each continent by year. 

```{r}
minLifeExp <- gapminder %>%
  group_by(year, continent) %>%
  summarise(min_lifeExp = min(lifeExp)) %>%
  spread(key = continent, value = min_lifeExp)
knitr::kable(minLifeExp)
```

Like in the plot in the first task, this form makes it easy to make plots that compare the different continents, since we have already divided them into seperate features. Lets make a line plot to get some variety:

```{r}
minLifeExp %>% 
  ggplot() +
  geom_line(aes(year, Africa), colour = 'red') +
  geom_line(aes(year, Asia), colour = 'green') +
  geom_line(aes(year, Europe), colour = 'purple') +
  geom_line(aes(year, Americas), colour = 'light blue') +
  geom_line(aes(year, Oceania), colour = 'orange') +
  labs(y = "Life Expectancy")
```

Here I'll create a new dataframe with some additional information about some selected countries, and join them to the year 2007 in gapminder, as that is roughly the same time as the olympic medal statistic is from:

```{r}
more_data <- data_frame(country = c('Norway', 'Canada', 'Sudan', 'China', 'Thailand'), highest_peak = c(2469, 5956, 3042, 8848, 2565), Olympic_goldMedals = c(188, 137, 0, 237, 9))

joined_data <- gapminder %>%
  filter(year == 2007) %>%
  right_join(more_data, by = 'country')
knitr::kable(joined_data)
```

Lets see if we can get different results using different joins:

```{r}
left_join <- gapminder %>%
  filter(year == 2007) %>%
  left_join(more_data, by = 'country')
knitr::kable(left_join)
```

So left_join will in this example keep all the rows in gapminder, and paste in the more_data dataset where it fits.


```{r}
full_join <- gapminder %>%
  filter(year == 2007) %>%
  full_join(more_data, by = 'country')
knitr::kable(full_join)
```

Actually, full_join does exactly the same as left_join on this data set. This is because all the countries present in more_data dataset, are also present in the original gapminder. Therefore, we should also see that inner_join will replicate the result of the right_join we started with:

```{r}
inner_join <- gapminder %>%
  filter(year == 2007) %>%
  inner_join(more_data, by = 'country')
knitr::kable(inner_join)
```


I can now check if there seems to be a correlation between the population in a country, and the number of olympic gold medals:

```{r}
ggplot(joined_data, aes(pop, Olympic_goldMedals)) +
  geom_point() +
  geom_line(colour = 'blue')
```

There definitely does not seem to be a trend here!

And since I am Norwegian I should probably brag a little bit, and show some statistics and plots of Olympic medals per population:

```{r}
joined_data %>%
  transmute(country, pop, Olympic_goldMedals, medalsPerPop = Olympic_goldMedals / pop) %>%
  ggplot(aes(country, medalsPerPop)) +
  geom_col(aes(fill = country))
```

Hehe, guess I might have to log-transform this data in the future, in order to at least be able to see the scores of the other countries ;)